---
title: "The Changing Student Concentration Composition at Harvard College"
author: "Neil Khurana"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(ggplot2)
library(janitor)
library(plotly)
```

# Overview

During November of each year, sophomores at the college are required to report their intended concentration to the Registrar. Here we have analyzed ten years of data to find out the concentrations that are growing, shrinking or rather steady. Our focus group for this analysis is divided into four categories: hard sciences, applied STEM, humanities, and social sciences. 

Major trends were noticable for each category. Applied science, technology, engineering and math concentrators have rapidly risen since 2008, whereas humanities concentrators are on the decline. Social science concentrations—Economics, Government, and Psychology—have continued to reamin the post popular and have seen no significant fluctuations. Same goes for the Hard Science concentrations, which includes chemistry and physics.

```{r concentrations, echo = FALSE}
concentrations <- read_xlsx("concentrations_h.xlsx") %>% 
  gather(Year, Count, 2:11) %>% 
  clean_names()
```
#### Hard Science Concentrators

```{r hard_sciences, echo = FALSE}
# Filtered to hard science concentrations only
hard_sciences <- concentrations %>% 
  filter(concentration %in% c("Molecular and Cellular Biology", "Chemistry", "Physics"))

# Stacked bar graph
hard_sciences_plot<- 
  ggplot(hard_sciences, aes(x = year, y = count, fill = concentration, label = count)) + 
  geom_bar(stat = "identity") +
  xlab("Year") +
  ylab("Number of Concentrators") +
  geom_text(size = 3, position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("limegreen", "chartreuse4", "darkgreen")); 
  plot(hard_sciences_plot)
```

#### Applied STEM Concentrators

```{R applied_stem, echo = FALSE}
# Data filtered to applied STEM concentrations only
applied_stem <- concentrations %>% 
  filter(concentration %in% c("Statistics", "Computer Science", "Applied Mathematics"))

# Stacked bar graph
applied_stem_plot<- 
  ggplot(applied_stem, aes(x = year, y = count, fill = concentration, label = count)) + 
  geom_bar(stat = "identity") + 
  xlab("Year") +
  ylab("Number of Concentrators") +
  geom_text(size = 3.5, position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("indianred1","red", "darkred"));
  plot(applied_stem_plot)
```

#### Humanities Concentrators

```{R humanities, echo = FALSE}
# Data filtered to humanities concentrations only
humanities <- concentrations %>% 
  filter(concentration %in% c("English", "Comparative Literature"))

# Stacked bar graph
humanities_plot<- 
  ggplot(humanities, aes(x = year, y = count, fill = concentration, label = count)) + 
  geom_bar(stat = "identity") +
  xlab("Year") +
  ylab("Number of Concentrators") +
  geom_text(size = 3, position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("plum3", "plum4"));
  plot(humanities_plot)
```

#### Social Sciences Concentrators

```{R social_sciences, echo = FALSE}
# Filtered to social science concentrations
social_sciences <- concentrations %>% 
  filter(concentration %in% c("Economics","Government", "Psychology"))

# Stacked bar graph
social_sciences_plot <- 
  ggplot(social_sciences, aes(x = year, y = count, fill = concentration, label = count)) + 
  geom_bar(stat = "identity") + 
  xlab("Year") +
  ylab("Number of Concentrators") +
  geom_text(size = 3, position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("deepskyblue","dodgerblue2", "dodgerblue4"));
  plot(social_sciences_plot)
```